---
title: NGS Assay and Project metadata 
format: html
date-modified: last-modified
date-format: long
date: 2023-11-30
bibliography: ../../resources/references.bib
summary: Examples of metadata files for experimental and computational purposes. 
---


:::{.callout-note title="Section Overview"}

⏰ **Time Estimation:** X minutes  

💬 **Learning Objectives:**   

1. Develop your metadata 
:::

You should consider revisiting these examples after completing [lesson 4](../04_metadata.qmd) in the course material. Please review these three tables containing pre-filled data fields for metadata, each serving distinct purposes: sample metadata, project metadata, and experimental metadata.

### Sample metadata fields

Some details might be specific to your samples. For example, which samples are treated, which are controlled, which tissue they come from, which cell type, the age, etc. Here is a list of possible metadata fields that you can use:

```{r}
#| echo: false
# Libraries
quiet <- function(x) { suppressMessages(suppressWarnings(x)) }
quiet(library(gt))
quiet(library(tidyverse))

# Load the data from CSV
data <- read_tsv('../assets/sample_metadata.tsv', show_col_types = FALSE)

# Print the data as a table
data %>% gt() |> tab_options(table.font.size = 12)
```

### Project metadata fields

Here you will find a table with possible metadata fields that you can use to annotate and track your `Project` folders:

```{r}
#| echo: false
# Libraries
quiet <- function(x) { suppressMessages(suppressWarnings(x)) }
quiet(library(gt))
quiet(library(tidyverse))

# Load the data from CSV
data <- read_tsv('../assets/project_metadata.tsv', show_col_types = FALSE)

# Print the data as a table
data %>% gt() |> tab_options(table.font.size = 12)
```

### Assay metadata fields

Here you will find a table with possible metadata fields that you can use to annotate and track your `Assay` folders:

```{r}
#| echo: false
# Libraries
quiet <- function(x) { suppressMessages(suppressWarnings(x)) }
quiet(library(gt))
quiet(library(tidyverse))

# Load the data from CSV
data <- read_tsv('../assets/assay_metadata.tsv', show_col_types = FALSE)

# Print the data as a table
data %>% gt()  |> tab_options(table.font.size = 12)
```

The metadata must include key details such as the project's short description, author information, creation date, experimental protocol, assay ID, assay type, platform utilized, library details, keywords, sample count, paired-end status, processor information, organism studied, sample origin, and file path. 

If you would create a database from the metadata files, your table should look like this (each row corresponding to one project):


```{r}
#| echo: false
# Libraries
quiet <- function(x) { suppressMessages(suppressWarnings(x)) }
quiet(library(gt))
quiet(library(tidyverse))

# Load the data from CSV
data <- read_tsv('../assets/database_example_20230810.tsv', show_col_types = FALSE)

# Print the data as a table
data %>% gt()  |> tab_options(table.font.size = 12)
```
